RadLib: a radiative heat transfer model library for CFD

### 2020-12-19 manuscript/manuscript.tex
updating manuscript.tex, somehow changes dont seem to have propgated, so Im trying to fix it

The Spectral Line Weighted-Sum-of-Gray Gases (SLW) model represents a family of global approaches to radiative heat transfer in high-temperature gases that also includes Absorption Distribution Function (ADF) and Full Spectrum $k$-distribution (FSK) models, all of which are based on the same fundamental principle in modeling the gas absorption spectrum \cite{Solovjov_2016}. In order to extend their spectral models from uniform conditions (isothermal, homogeneous gases) to nonuniform conditions (non-isothermal, non-homogeneous gases), these models take a reference approach in which local gas states are corrected relative to a reference state. Reference approaches, however, generally lack consistent reference states and can yield significant errors in cases with large spatial temperature gradients \cite{Solovjov_2017}. The Rank Correlated SLW (RCSLW) model is a unique extension of the generalized SLW model that does not require a specified gas reference state and preserves the emission term of the spectrally integrated RTE. 
%
% TODO: what does this last phrase mean: preserves the emission...?
%
Recent comparison of advanced SLW modeling approaches revealed that ``the Rank Correlated SLW model is the most robust of all models, and demonstrates that it can achieve accurate solutions with as few as 3–5 gray gases" \citep{Badger_2019}. A brief overview of the general SLW and RCSLW models will be given here; detailed discussion can be found in the literature \cite{Solovjov_2000, Solovjov_2001, Solovjov_2008, Solovjov_2011, Solovjov_2014, Solovjov_2016, Solovjov_2017, Webb_2018}.

SLW models are similar to WSGG models but instead of integrating over simple sequential $\eta$ bands, spectral integration is done over wavenumbers in a given range of absorption cross section. So, in Eq.~(\ref{e:RTE_wsgg}), gas $j$ would refer to wavenumbers $\eta\in\Delta_j$ with $\Delta_j=\{\eta:\,\tilde{C}_{j-1}<C_\eta<\tilde{C}_j\}$, using Solovjov's notation, where $C$ denotes absorption cross-section (where $\kappa=CN$, $N$ is molar density of the radiatively participating gas), and $\tilde{C}$ are supplemental cross sections defining interfaces between adjacent cross-section bands.

If $\tilde{C}_j$ are known for all $j$, then $\kappa_j = N\sqrt{C_{j-1}C_j}$. The weights $a_j$ are computed using the absorption line blackbody distribution function (ALBDF), denoted $F$, which gives the fraction of total blackbody emission from $\eta$ corresponding to absorbtion cross sections less than $C$: $F$ is a monotonically-increasing function of $C$, and depends on the thermodynamic state $\phi$, and the black temperature $T_b$ defining the Planck function: $F = F(C,\phi,T_b)$. The $a_j$ are then computed as
%
\begin{linenomath}
\begin{equation}
a_j = F(\tilde{C}_j,\phi, T) - F(\tilde{C}_{j-1}, \phi,T).
\end{equation}
\end{linenomath}
%

Equation~(\ref{e:RTE_wsgg}) assumes the $\Delta_j$ intervals are independent of spatial position, but care must be taken since $C(\eta)$ depends on the thermodynamic state, which does vary spatially. Hence, the fixed $\Delta_j$ imply a corresponding variation in $C_j$ and $\tilde{C}_j$.

>>>>

The Spectral Line Weighted-Sum-of-Gray Gases (SLW) model represents a family of global approaches to radiative heat transfer in high-temperature gases that also includes Absorption Distribution Function (ADF) and Full Spectrum $k$-distribution (FSK) models, all of which are based on the same fundamental principle in modeling the gas absorption spectrum \cite{Solovjov_2016}. In order to extend their spectral models from uniform conditions (isothermal, homogeneous gases) to nonuniform conditions (non-isothermal, non-homogeneous gases), these models take a reference approach in which local gas states are corrected relative to a reference state. Reference approaches, however, generally lack consistent reference states and can yield significant errors in cases with large spatial temperature gradients \cite{Solovjov_2017}. The Rank Correlated SLW (RCSLW) model is a unique extension of the generalized SLW model that does not require a specified gas reference state and preserves the emission term of the spectrally integrated RTE. 
% TODO: what does this last phrase mean: preserves the emission...?
%
Recent comparison of advanced SLW modeling approaches revealed that ``the Rank Correlated SLW model is the most robust of all models, and demonstrates that it can achieve accurate solutions with as few as 3–5 gray gases" \citep{Badger_2019}. A brief overview of the RCSLW model is given here; detailed discussion can be found in the literature \cite{Solovjov_2000, Solovjov_2001, Solovjov_2008, Solovjov_2011, Solovjov_2014, Solovjov_2016, Solovjov_2017, Webb_2018}. 
Solovjov et al.~\cite{Solovjov_2017} present a number of methods within a generalized framework, with method I.2.2 giving the best results. That is the method implemented and summarized here.
%

SLW models are similar to WSGG models but instead of integrating over a given range of $\eta$, with sequential $\eta$ bands defining gray gases $j$, spectral integration is done over wavenumbers in a given range of absorption cross section, with sequential bands of cross-section defining the gray gases $j$. So, in Eq.~(\ref{e:RTE_wsgg}), gas $j$ would refer to wavenumbers $\eta\in\Delta_j$ with $\Delta_j=\{\eta:\,\tilde{C}_{j-1}<C_\eta<\tilde{C}_j\}$, using Solovjov's notation, where $C$ denotes absorption cross-section (where $\kappa=CN$ and $N$ is molar density of the radiatively participating gas), and $\tilde{C}$ are supplemental cross sections defining interfaces between adjacent cross-section bands. This ensures that the gas is in fact gray in a given band, by definition, at least to within the width $\tilde{C}_j-\tilde{C}_{j-1}$. In contrast, in traditional WSGG approaches, the gases are only conceptually gray, and the absorption coefficients and weights are determined empirically \cite{Badger_2019}.

The gas weights $a_j$, are the fraction of the blackbody emission from $\eta$ corresponding to a given cross-section band. This energy is computed using the absorption line blackbody distribution function (ALBDF) \cite{Pearson_2014}, which gives the fraction of total blackbody emission from $\eta$ corresponding to absorption cross sections less than $C$: 
%
\begin{linenomath}
    \begin{equation} \label{e:albdf}
    F(C,\phi,T) = \frac{1}{\sigma T^4}\int_{\{\eta:\,C(\eta,\phi)<C\}}E_b(\eta,T)d\eta,
\end{equation}
\end{linenomath}
%
so that
%
\begin{linenomath}
\begin{equation}
a_j = F(\tilde{C}_j,\phi, T) - F(\tilde{C}_{j-1}, \phi,T).
\end{equation}
\end{linenomath}
%
%todo: deal with T versus Tb in e:albdf and previous eqn.
Note that $F$ varies between zero and one, and is monotonic in $C$, allowing calculation of the inverse $C(F,\phi,T_b)$.

Equation~(\ref{e:RTE_wsgg}) assumes the sets $\Delta_j$ are independent of spatial position, but care must be taken since C depends on the thermodynamic state (denoted by vector $\phi$ with nominal components $T$, $P$, and $y_i$), which does vary spatially. That is, $C=C(\eta,\phi)$. Hence, the fixed $\Delta_j$ imply a corresponding variation in $C_j$ and $\tilde{C}_j$ from one spatial location to the next.

The RCSLW model uses the notion of \emph{rank correlation} to account for this variation in the $C_j$, which in turn define the absorption coefficients $\kappa_j$. Solovjov defines two spectra arising from thermodynamic states $\phi_1$ and $\phi_2$ to be rank correlated if, for an arbirary wavenumber $\hat{\eta}$, the wave number sets $\Delta_1$ and $\Delta_2$ are equal, where $\Delta_1=\{\eta: C(\eta,\phi_1)<C(\hat{\eta},\phi_1)\}$, and $\Delta_2=\{\eta: C(\eta,\phi_2)<C(\hat{\eta},\phi_2)\}$. 
Note in Eq.~(\ref{e:albdf}), that $\phi$ determines $C$, which then determines the set of $\eta$ integrated over. But for a \emph{given} set $\eta$, e.g., $\Delta_1$, $F$ can be considered as $F(C,T)$. Hence, if $\Delta_1=\Delta_2$, then at a fixed reference black temperature $T_{ref}$, we have $F(C_1,\phi_1,T_{ref})=F(C_2,\phi_2,T_{ref})$. One can then compute the inverse to find $C_1$ at $\phi_1$ and $C_2$ at $\phi_2$ (i.e., values at two spatial positions) corresponding to the same wavenumber interval.

In practice, for a given problem, a grid of $\tilde{F}_j$ points are specified, with $F_j$ points between $\tilde{F}_j$ and $\tilde{F}_{j-1}$. The corresponding $\tilde{C}_j$ and $C_j$ values at any spatial position can then be found by inversion: $C_j = C(F_j,\phi,T_{ref})$, and similarly for $\tilde{C}_j$ in terms of $\tilde{F}_j$. $\kappa_j$ and $a_j$ can then be computed as noted above.
%TODO verify this in terms of the quadrature.
